---
title: "Reproducible_Research"
author: "Chuanneng Sun"
date: "November 28, 2017"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### First, read the data
```{r}
data <- read.csv('./activity.csv')
```

Let's see what the data looks like.
```{r}
head(data)
```
So there are a lot of missing values in "Step" attribute. So be careful when making plots.

### Make some plots of the data
```{r}
library(lattice)
library(dplyr)
library(ggplot2)
hist(data$steps, xlab = 'Steps', main = 'Frequency of Steps')
```
So we can see a basic distribution of steps.

### Then we should calculate the mean and median number of steps
```{r}
data %>%
summarise(mean=mean(steps, na.rm = T), median=median(steps, na.rm = T)) %>%
print
```
### Next, we ought to look at the time series of steps
```{r}
group_data <- group_by(data, date)
group_data <- summarise(group_data, mean=mean(steps))
ggplot(data=group_data, aes(x=date, y=mean)) +
    geom_line() + geom_point()
```

### Next, let's find out which interval contains the most steps
```{r}
data[which.max(data$steps), ]
```

So on 2012-11-27, steps come up to 806 within 5 minutes. Amazing!

### Next let's find out a way to imput the missing data.
```{r}
missing_data <- data[!complete.cases(data$steps), ]
ggplot(data=missing_data, aes(x=date)) + stat_count(aes(fill = date))
```

What's this? Seems that data of these days have been missing.
```{r}
missing_data %>%
    group_by(date) %>%
    summarise(n=n()) %>%
    print
```
Huh!, so data of these days are really missing.
I tend to omit these missing values.
```{r}
good_data <- data[complete.cases(data$steps), ]
```

### So lets plot the total number of steps
```{r}
good_data %>%
    group_by(date) %>%
    summarise(sum=sum(steps)) %>%
    ggplot(aes(x=date, y=sum)) + geom_bar(stat = "identity") +
    theme(axis.text.x = element_text(angle = 90))
```

### Finally, we should compare number of steps in weekdays with weekends
```{r}
good_data$date <- as.Date(as.character(good_data$date))
good_data$weekday <- weekdays(good_data$date)
head(good_data)
```
So here we can see the weekdays of each day.
```{r}
good_data %>%
    group_by(weekday) %>%
    summarise(avg=mean(steps)) %>%
    ggplot(aes(x=weekday, y=avg)) + geom_bar(stat='identity', aes(fill=weekday))
```
To further explore the difference between weekdays and weekends, we classify the weekdays into 'weekday' and 'weekend'.
```{r}
good_data$type <- 'NA'
good_data[good_data$weekday %in% c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday'), ]$type <- 'weekday'
good_data[good_data$weekday %in% c('Saturday', 'Sunday'), ]$type <- 'weekend'
head(good_data)
```
Let's plot it!!
```{r}
good_data %>%
    group_by(type) %>%
    summarise(avg=mean(steps)) %>%
    ggplot(aes(x=type, y=avg, fill=type)) + geom_bar(stat='identity')
```

It's clear to see that steps on weekends are more than on weekdays.