---
title: "Reproducible_Research"
author: "Chuanneng Sun"
date: "November 28, 2017"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lattice)
library(dplyr)
library(ggplot2)
library(Hmisc)
```

### First, read the data
```{r}
data <- read.csv('./activity.csv')
```

Let's see what the data looks like.
```{r}
head(data)
```
So there are a lot of missing values in "Step" attribute. So be careful when making plots.


### Total number of steps each day
```{r}
daySteps <- tapply(data$steps, data$date, sum, na.rm=TRUE)
qplot(daySteps, xlab='Steps per day', ylab='Frequency')
```
We can get a brief impression about the data.

### Mean and median steps each day
```{r}
dayStepsMean <- mean(daySteps)
dayStepsMedian <- median(daySteps)
print(c(dayStepsMean, dayStepsMedian))
```
So we can see clearly hear that the mean and median of the data is 9354.23 and 10395.00.


### Average daily activity
```{r}
aggSteps <- aggregate(x=list(steps=data$steps),by=list(interval=data$interval), FUN=mean, na.rm=TRUE)
head(aggSteps)
```

### Next, we ought to make a time series of steps using the processed data above
```{r}
ggplot(aggSteps, aes(interval, steps)) + geom_line()
```

### Next, let's find out which interval contains the most steps
```{r}
mostStep <- which.max(aggSteps$steps)
mostInterval <- aggSteps[mostStep, 'interval']
mostInterval
```

So on it is #835 interval which contains the most steps.

### Next let's find out a way to imput the missing data.
```{r}
missing_data <- data[!complete.cases(data$steps), ]
ggplot(data=missing_data, aes(x=date)) + stat_count(aes(fill = date)) +
theme(axis.text.x = element_text(angle = 45))
```

What's this? Seems that data of these days have been missing.
```{r}
missing_data %>%
    group_by(date) %>%
    summarise(n=n()) %>%
    print
```
Huh!, so data of these days are really missing.
Use mean value of the dataset to impute the missing values.
```{r}
data$steps <- impute(data$steps, fun=mean)
head(data)
```

### Let's plot the imputed histogram of the data.
```{r}
dayStepsImp <- tapply(data$steps, data$date, sum)
qplot(dayStepsImp)
```

### Finally, we should compare number of steps in weekdays with weekends
```{r}
data$date <- as.Date(as.character(data$date))
data$weekday <- weekdays(data$date)
head(data)
```
So here we can see the weekdays of each day.
```{r}
data %>%
    group_by(weekday) %>%
    summarise(avg=mean(steps)) %>%
    ggplot(aes(x=weekday, y=avg)) + geom_bar(stat='identity', aes(fill=weekday))
```
To further explore the difference between weekdays and weekends, we classify the weekdays into 'weekday' and 'weekend'.
```{r}
data$type <- 'NA'
data[data$weekday %in% c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday'), ]$type <- 'weekday'
data[data$weekday %in% c('Saturday', 'Sunday'), ]$type <- 'weekend'
head(data)
```


```{r}
data %>%
    group_by(type) %>%
    summarise(avg=mean(steps)) %>%
    ggplot(aes(x=type, y=avg, fill=type)) + geom_bar(stat='identity')
```

It's clear to see that steps on weekends are more than on weekdays.